---
title: 'Final Study Data Analysis'
author: "April Kim, Jennifer Podracky, Saurav Datta"
output:
  pdf_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(data.table)
library(lmtest)
library(pwr)
library(lsr)
```


*Read in data and reformat*
```{r}
d2 <- read.csv("241 Participant List - Final Study Results - 20181215.csv", 
                    stringsAsFactors = F, na.strings=c("","NA"))

d2$Q7 <- as.numeric(gsub("\\,", "", d2$Q7))
d2$Q11 <- as.numeric(gsub("\\,", "", d2$Q11))
d2$Q15 <- as.numeric(gsub("\\,", "", d2$Q15))

# Not applicable = 0
# Through digital means = 1
# In person = 2

d2[d2$Q6 %like% "Not applicable", ]$Q6 <- 0
d2[d2$Q10 %like% "Not applicable", ]$Q10 <- 0
d2[d2$Q14 %like% "Not applicable", ]$Q14 <- 0

d2[d2$Q6 %like% "In person", ]$Q6 <- 2
d2[d2$Q10 %like% "In person", ]$Q10 <- 2
d2[d2$Q14 %like% "In person", ]$Q14 <- 2

d2[d2$Q6 %like% "Through digital means", ]$Q6 <- 1
d2[d2$Q10 %like% "Through digital means", ]$Q10 <- 1
d2[d2$Q14 %like% "Through digital means", ]$Q14 <- 1

# names(d2) <- c("name", "email", "know_us", "day1_treatment", "step_day1", "day2_treatment", "step_day2", "day3_treatment", "step_day3", "treatment_seq")

names(d2) <- c("user_id", "treatment_seq", "location_lat", "location_long", "name", "email", "gender", "gender_code", "age_range", "age_range_code", "location", "lives_with_others", "lives_with_others_code", "submitted_response", "noncompliance_ind", "know_us",
               "day1_num_persons", "day1_treatment", "day1_steps", "day1_upload_id", "day1_upload_filename", "day1_upload_size", "day_1_upload_filetype", 
               "day2_num_persons", "day2_treatment", "day2_steps", "day2_upload_id", "day2_upload_filename", "day2_upload_size", "day_2_upload_filetype",
               "day3_num_persons", "day3_treatment", "day3_steps", "day3_upload_id", "day3_upload_filename", "day3_upload_size", "day3_upload_filetype")

d2$know_us[d2$know_us == "Yes"] <- 1
d2$know_us[d2$know_us == "No"] <- 0
d2$know_us[is.na(d2$know_us)] <- 2
d2$know_us <- as.numeric(d2$know_us)


# covariate balance

bal.tab(treatment_seq ~ gender_code + age_range_code + lives_with_others_code + know_us + location_lat + location_long, data = d2,
         method = "weighting")
```



```{r}
# for treatment received on what day below
# 1 = day1
# 2 = day2
# 3 = day3

d2$treatment1_day_received <- NA
d2$treatment2_day_received <- NA
d2$treatment3_day_received <- NA

d2[d2$treatment_seq == 1,]$treatment1_day_received <- 1
d2[d2$treatment_seq == 1,]$treatment2_day_received <- 2
d2[d2$treatment_seq == 1,]$treatment3_day_received <- 3

d2[d2$treatment_seq == 2,]$treatment1_day_received <- 1
d2[d2$treatment_seq == 2,]$treatment2_day_received <- 3
d2[d2$treatment_seq == 2,]$treatment3_day_received <- 2

d2[d2$treatment_seq == 3,]$treatment1_day_received <- 2
d2[d2$treatment_seq == 3,]$treatment2_day_received <- 1
d2[d2$treatment_seq == 3,]$treatment3_day_received <- 3

d2[d2$treatment_seq == 4,]$treatment1_day_received <- 2
d2[d2$treatment_seq == 4,]$treatment2_day_received <- 3
d2[d2$treatment_seq == 4,]$treatment3_day_received <- 1

d2[d2$treatment_seq == 5,]$treatment1_day_received <- 3
d2[d2$treatment_seq == 5,]$treatment2_day_received <- 1
d2[d2$treatment_seq == 5,]$treatment3_day_received <- 2

d2[d2$treatment_seq == 6,]$treatment1_day_received <- 3
d2[d2$treatment_seq == 6,]$treatment2_day_received <- 2
d2[d2$treatment_seq == 6,]$treatment3_day_received <- 1
```

*Checking for ordering/priming effect*
For this part, we're interested in seeing if receiving certain treatment on what day has any effect on the outcome, so the outcome will be whether or not the subject had more than 5000 steps on the third day. Analysis will be limited to those who actually followed their treatment assignment.

```{r}
d2_a <- d2
d2_a$outcome <- ifelse(d2_a$day3_steps > 5000, 1, 0)

# limit to those who followed directions
d2_a <- rbind(d2_a[d2_a$treatment_seq == 1 & d2_a$day1_treatment == 0 & d2_a$day2_treatment == 1 & d2_a$day3_treatment == 2,],
              d2_a[d2_a$treatment_seq == 2 & d2_a$day1_treatment == 0 & d2_a$day2_treatment == 2 & d2_a$day3_treatment == 1,],
              d2_a[d2_a$treatment_seq == 3 & d2_a$day1_treatment == 1 & d2_a$day2_treatment == 0 & d2_a$day3_treatment == 2,],
              d2_a[d2_a$treatment_seq == 4 & d2_a$day1_treatment == 1 & d2_a$day2_treatment == 2 & d2_a$day3_treatment == 0,],
              d2_a[d2_a$treatment_seq == 5 & d2_a$day1_treatment == 2 & d2_a$day2_treatment == 0 & d2_a$day3_treatment == 1,],
              d2_a[d2_a$treatment_seq == 6 & d2_a$day1_treatment == 2 & d2_a$day2_treatment == 1 & d2_a$day3_treatment == 0,])

# interested in interaction between treatment and day
m_a <- lm(outcome ~ day1_treatment:treatment1_day_received + day2_treatment:treatment2_day_received + day3_treatment:treatment3_day_received +
             day1_steps + day2_steps, data = d2_a)
m_a
coeftest(m_a)
```
Sequence of treatment isn't likely to impact outcome.



T-test and power calculations
```{r}
# get data frame in workable format
d <- rbindlist(list(d2_a[1:nrow(d2_a),c("day1_treatment","day1_steps")],
                    d2_a[1:nrow(d2_a),c("day2_treatment","day2_steps")],
                    d2_a[1:nrow(d2_a),c("day3_treatment","day3_steps")]))
names(d) <- c("treatment","steps")
d$outcome <- ifelse(d$steps > 5000, 1, 0)

###  Control vs digital 
t.test(d[treatment == 0]$outcome, d[treatment == 1]$outcome, paired = T)
cohensD(d[treatment == 0]$outcome, d[treatment == 1]$outcome, method = "paired")

### Control vs in person 
t.test(d[treatment == 0]$outcome, d[treatment == 2]$outcome, paired = T)
cohensD(d[treatment == 0]$outcome, d[treatment == 2]$outcome, method = "paired")

### In person vs digital 
t.test(d[treatment == 2]$outcome, d[treatment == 1]$outcome, paired = T)
```
no significant difference in both comparisons

covariate balance
```{r}

```